Concurrent Programming in Harmony

### 2021-08-01 book/paper.tex
Latest

Harmony provides support for testing linearizability,
but requires that the programmer
identifies what are known as \emph{linearization points}
in the implementation that indicate exactly \emph{which} sequential
execution the concurrent execution must align with.
\autoref{fig:queuelin} is a copy of \autoref{fig:queue} extended with
linearization points.
For each operation (\texttt{get} and \texttt{put}), the corresponding
linearization point must occur somewhere between acquiring and releasing
the lock.
%
Each linearization point execution is assigned a logical timestamp.
Logical timestamps are numbered $0, 1, ...$
To do so, we have added a counter (\texttt{time}) to the \texttt{Queue}.
Method \texttt{\_linpoint} saves the current counter in
\textbf{this}.\texttt{qtime} and increments the counter.
The \textbf{this} dictionary maintains \emph{thread-local state} associated
with the thread (\autoref{ch:harmonymachine})---it contains variables that
can be accessed by any method in the thread.

@@

\begin{figure}
\begin{code}
\harmonysource{abd1}
\end{code}
\caption{\harmonylink{code/abd.hny} An implementation of a replicated atomic read/write register.}
\label{fig:abd1}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{abd2}
\end{code}
\caption{\harmonylink{code/abd.hny} Testing the code in \autoref{fig:abd1}.}
\label{fig:abd2}
\end{figure}

>>>>

\index{logical timestamp}%
Harmony provides support for testing linearizability,
but requires that the programmer
identifies what are known as \emph{linearization points}
in the implementation that indicate exactly \emph{which} sequential
execution the concurrent execution must align with.
\autoref{fig:queuelin} is a copy of \autoref{fig:queue} extended with
linearization points.
For each operation (\texttt{get} and \texttt{put}), the corresponding
linearization point must occur somewhere between acquiring and releasing
the lock.
%
Each linearization point execution is assigned a logical timestamp.
Logical timestamps are numbered $0, 1, ...$
To do so, we have added a counter (\texttt{time}) to the \texttt{Queue}.
Method \texttt{\_linpoint} saves the current counter in
\textbf{this}.\texttt{qtime} and increments the counter.
The \textbf{this} dictionary maintains \emph{thread-local state} associated
with the thread (\autoref{ch:harmonymachine})---it contains variables that
can be accessed by any method in the thread.

@@

\begin{figure}
\begin{code}
\harmonysource{abd}
\end{code}
\caption{\harmonylink{code/abd.hny} An implementation of a replicated atomic read/write register.}
\label{fig:abd}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{abdtest}
\end{code}
\caption{\harmonylink{code/abdtest.hny} Testing the code in \autoref{fig:abd}.}
\label{fig:abdtest}
\end{figure}

A \emph{register} is an object that you can read or write.  In a distributed
system, it can be useful to have shared registers.  A simple shared register
would be maintained by a server, and clients can read or write the
shared register by exchanging messages with the server.
We call two operations such that one does not finish before the other starts
\emph{concurrent}.
Since messages are delivered one at a time to the server,
concurrent operations on the shared register appear atomic.
In particular, we have the following three desirable properties:
\begin{enumerate}
\item All write operations are ordered;
\item A read operation returns either the last value written or the value of
a concurrent write operation.
\item If read operation $r_1$ finishes before read operation $r_2$ starts,
then $r_2$ cannot return a value that is older than the value returned
by $r_1$.
\end{enumerate}

\index{single point of failure}%
Unfortunately, however, the server is a \emph{single point of failure}:
if it fails, all its clients suffer.
We would therefore like to find a solution that can survive the crash
of a server.
In this chapter, we will \emph{replicate} the object: maintain multiple
copies.
One could, for example, imaging that clients write to all copies and read
from any single one.
While this solves the single-point-of-failure problem, we lose all the
nice properties above.  For one, it is not guaranteed that all servers
see all write operations in the same order.

\index{logical timestamp}%
We present a protocol here that is based on the work by
Hagit Attiya, Amotz Bar-Noy, and Danny Dolev~\cite{ABD95}.
In order to tolerate \texttt{N} failures,
it uses $\mathtt{N} = 2\mathtt{F} + 1$ servers.
Each server maintains not only the value of the object, but also a
\emph{logical timestamp} (see also \autoref{ch:testing}).
All write operation will be ordered by a unique timestamp.
